Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-09-03 draft.tex
numbers now properly delimited by points

We only display variables with an error rate below $0,5$, results above this threshold are inaccurate.

@@

\subsubsection{All Variables}
%%
The vast majority of parameters can not be described well using the classifier.
%%
This is reflected in the large amount of variables classified with an error rate above $0,6$.
%%
%The first impression of the resulting scatter plots quickly shows a very high amount of variables above an error rate of $0,6$.
%%

% None of the pain indicators can be reliably described using the image-based parameter.
None of the pain indicators can be described reliably using the image-based parameters.
%%
The only variable reliably classified in this group is gender. % with an error rate of 31,5\% using a decision tree of seven rules.
%
It can be discriminated with an error rate of $0,31$ using 7 rules and incorporates only curvature and curvature angle variables.
%%
The distinctness lies in the average difference in body size between male and female subjects.
%
%It mostly incorporates \emph{mean curvature}, \emph{sagittal curvature angle} und sagittal \emph{mean curvature}.
%The perfect categorization of the \emph{dialysis} parameter was conspicuous.
%%
%This is due to the subject distribution among the manifestation.
%%
%Only two subjects receive a dialysis.
%%
%The good discrimination is not meaningful.
%%
\emph{Medication for high blood pressure} is classified for 1,058 subjects with an error rate of $0,47$ solely based on coronal mean curvature.
%%
%1,058 subjects are classified with an error rate of $0,47$ solely based on coronal mean curvature.
%%
Almost all subjects who are medicated ($796/1,058$) were correctly classified, the vast majority of non-medicated subjects ($262/1,058$) are false-positive classified.
%%
Therefore the classifier is not as useful.
%%This is however due to the high amount of people, who are medicated (68\%) and who are almost all correctly class
%%
%Other features well classified include marital status, foot pain and number of born children.
%%
The four body size groups could be characterized with an error rate $0,48$, but the decision tree comprises 71 rules and imposes overfitting.

The analysis of the dummy variable yields similar results like the blood pressure medication.
%%
Variables, such subjects sized 139-164 cm, between 64 and 90 years of age or nutrition related parameters are dominantly populated by one manifestation.
%%
The classifier neglects the other group and yields a error rate below $0,5$.
%The dummy variable analysis yields fewer results.
%%
%It shows that subjects between 64 and 90 years of age can be detected with 23\% error rate by using the curvature angle and coronal mean curvature.
%%
%Subjects with body size between 177 cm and 202 cm can classified with an error rate of 22,1\%. solely using mean curvature.
%%
\subsubsection{Gender Groups}
%%
Classification using groups divided by gender do not produce satisfying results.
%%
Only hypothyroidism could be described for male subjects with an error rate of $0,24$ for 110 subjects using the \emph{mean curvature} and \emph{curvature angle}.
%%
Since there are only 30 male subjects diagnosed with hypothyroidism, the statistical power of the result is reduced.
%%
The dummy variable analysis showed that female subjects of 139-164 cm body size could be discriminated using the mean curvature and curvature angle, with an error of $0,38$.

@@

\subsubsection{Size Groups}
%%
Many previously described results are confounded with subject size.
%%
Differences between male and female subjects are explained by to the average body size difference.
%%
For example, large subjects are already characterized by their rather straight spine.
%
The question is, whether the inter-group spine-variability parameter is enough for predicting other parameter or not.
%%
Dividing subjects into body size groups potentially highlights classifications not confounded by body size.
%%
\paragraph{Large Decision Trees. }
Back pain-associated variables can be explained for various size-groups, but we could not extract universal rules.
%%
Radiating back pain could be described with error rate of $0.39$ using 23 rules for subjects between $171-177~cm$ body size using \emph{torsion} and \emph{mean curvature}.
%%
For subjects sized $177-202~cm$ the accuracy drops to $0.47$ using 20  decision rules.
%%
There are several decision trees for laboratory values, such as alanine aminotransferase value (relevant for diagnosis of lover or gallbladder illness) in the blood can be described with an error rate of $0.4$ ($139-164~cm$) to $0.36$ ($164 - 171~cm$).
%%
Similar values can be observed for cholesterol or age.
%%
Due to the large decision trees, these results are not usable and impose overfitting to the data.
%%
\paragraph{Small Decision Trees. }
%%
The dummy variables show several variables described using only one decision rule with accuracy between $0,42$ to $0,47$.
%%
Most of these variables have a dominant manifestation and the classifier shows a low detection precision for the second manifestation.
%%
These variables include nutrition parameter, thyroid disorder and social problems induced by back pain.

>>>>

We only display variables with an error rate below $0.5$, results above this threshold are inaccurate.

@@

\subsubsection{All Variables}
%%
The vast majority of parameters can not be described well using the classifier.
%%
This is reflected in the large amount of variables classified with an error rate above $0.6$.
%%
%The first impression of the resulting scatter plots quickly shows a very high amount of variables above an error rate of $0.6$.
%%

% None of the pain indicators can be reliably described using the image-based parameter.
None of the pain indicators can be described reliably using the image-based parameters.
%%
The only variable reliably classified in this group is gender. % with an error rate of 31.5\% using a decision tree of seven rules.
%
It can be discriminated with an error rate of $0.31$ using 7 rules and incorporates only curvature and curvature angle variables.
%%
The distinctness lies in the average difference in body size between male and female subjects.
%
%It mostly incorporates \emph{mean curvature}, \emph{sagittal curvature angle} und sagittal \emph{mean curvature}.
%The perfect categorization of the \emph{dialysis} parameter was conspicuous.
%%
%This is due to the subject distribution among the manifestation.
%%
%Only two subjects receive a dialysis.
%%
%The good discrimination is not meaningful.
%%
\emph{Medication for high blood pressure} is classified for 1,058 subjects with an error rate of $0.47$ solely based on coronal mean curvature.
%%
%1,058 subjects are classified with an error rate of $0,47$ solely based on coronal mean curvature.
%%
Almost all subjects who are medicated ($796/1,058$) were correctly classified, the vast majority of non-medicated subjects ($262/1,058$) are false-positive classified.
%%
Therefore the classifier is not as useful.
%%This is however due to the high amount of people, who are medicated (68\%) and who are almost all correctly class
%%
%Other features well classified include marital status, foot pain and number of born children.
%%
The four body size groups could be characterized with an error rate $0.48$, but the decision tree comprises 71 rules and imposes overfitting.

The analysis of the dummy variable yields similar results like the blood pressure medication.
%%
Variables, such subjects sized 139-164 cm, between 64 and 90 years of age or nutrition related parameters are dominantly populated by one manifestation.
%%
The classifier neglects the other group and yields a error rate below $0.5$.
%The dummy variable analysis yields fewer results.
%%
%It shows that subjects between 64 and 90 years of age can be detected with 23\% error rate by using the curvature angle and coronal mean curvature.
%%
%Subjects with body size between 177 cm and 202 cm can classified with an error rate of 22,1\%. solely using mean curvature.
%%
\subsubsection{Gender Groups}
%%
Classification using groups divided by gender do not produce satisfying results.
%%
Only hypothyroidism could be described for male subjects with an error rate of $0.24$ for 110 subjects using the \emph{mean curvature} and \emph{curvature angle}.
%%
Since there are only 30 male subjects diagnosed with hypothyroidism, the statistical power of the result is reduced.
%%
The dummy variable analysis showed that female subjects of 139-164 cm body size could be discriminated using the mean curvature and curvature angle, with an error of $0.38$.

@@

\subsubsection{Size Groups}
%%
Many previously described results are confounded with subject size.
%%
Differences between male and female subjects are explained by to the average body size difference.
%%
For example, large subjects are already characterized by their rather straight spine.
%
The question is, whether the inter-group spine-variability parameter is enough for predicting other parameter or not.
%%
Dividing subjects into body size groups potentially highlights classifications not confounded by body size.
%%
\paragraph{Large Decision Trees. }
Back pain-associated variables can be explained for various size-groups, but we could not extract universal rules.
%%
Radiating back pain could be described with error rate of $0.39$ using 23 rules for subjects between $171-177~cm$ body size using \emph{torsion} and \emph{mean curvature}.
%%
For subjects sized $177-202~cm$ the accuracy drops to $0.47$ using 20  decision rules.
%%
There are several decision trees for laboratory values, such as alanine aminotransferase value (relevant for diagnosis of lover or gallbladder illness) in the blood can be described with an error rate of $0.4$ ($139-164~cm$) to $0.36$ ($164 - 171~cm$).
%%
Similar values can be observed for cholesterol or age.
%%
Due to the large decision trees, these results are not usable and impose overfitting to the data.
%%
\paragraph{Small Decision Trees. }
%%
The dummy variables show several variables described using only one decision rule with accuracy between $0.42$ to $0.47$.
%%
Most of these variables have a dominant manifestation and the classifier shows a low detection precision for the second manifestation.
%%
These variables include nutrition parameter, thyroid disorder and social problems induced by back pain.
